### 2023-02-27 paper.tex
add some methods

\documentclass[]{article}
\usepackage[super,numbers]{natbib}

\usepackage[margin=1in]{geometry}
\usepackage{url}
\usepackage{graphicx}
\usepackage{color}
\usepackage{booktabs}

\usepackage{authblk}

\usepackage{lineno}
%\linenumbers
\renewcommand{\Affilfont}{\normalsize}
\renewcommand{\Authfont}{\normalsize}

% local definitions
\newcommand{\comment}[1]{{\textcolor{red}{Co-author: #1}}}
\newcommand{\sgcomment}[1]{{\textcolor{red}{SG: #1}}}
\newcommand{\bmhcomment}[1]{{\textcolor{blue}{BMH: #1}}}
\newcommand{\aprcomment}[1]{{\textcolor{magenta}{APR: #1}}}
\newcommand{\tdwcomment}[1]{{\textcolor{cyan}{TDW: #1}}}

% cross-reference with supplement
\usepackage{xr}
\externaldocument{supplement}

\begin{document}

\title{A weakly structured stem for human origins in Africa}
\author[1]{Aaron~P.~Ragsdale}
\author[2]{Timothy~D.~Weaver}
\author[3]{Elizabeth~G.~Atkinson}
\author[4]{Eileen~Hoal}
\author[4]{Marlo~M\"{o}ller}
\author[2,5,$\dag$,*]{Brenna~M.~Henn}
\author[6,$\dag$,**]{Simon~Gravel}
\affil[1]{Department of Integrative Biology, University of Wisconsin--Madison, WI, USA}
\affil[2]{Department of Anthropology, University of California, Davis, Davis, CA, USA}
\affil[3]{Department of Molecular and Human Genetics, Baylor College of Medicine, Houston, TX, USA}
\affil[4]{DSI-NRF Centre of Excellence for Biomedical Tuberculosis Research; South African Medical Research Council Centre for Tuberculosis Research; Division of Molecular Biology and Human Genetics, Faculty of Medicine and Health Sciences, Stellenbosch University, Cape Town, South Africa}
\affil[5]{UC Davis Genome Center, University of California, Davis, Davis, CA, USA}
\affil[6]{Department of Human Genetics, McGill University, Montreal, QC, Canada}
\affil[$\dag$]{Co-Corresponding Authors}
\affil[*]{bmhenn@ucdavis.edu}
\affil[**]{simon.gravel@mcgill.ca}
\date{\normalsize \today}
\maketitle

\begin{abstract}

    Despite broad agreement that \emph{Homo sapiens} originated within Africa,
    considerable uncertainty surrounds specific models of divergence and
    migration across the continent\citep{Henn2018-rf}. Progress is hampered by limited fossil and
    genomic data, as well as variability in prior divergence time estimates\citep{Henn2018-rf}.
    To discriminate among such models, we consider linkage disequilibrium and diversity-based 
    statistics, optimized for rapid, complex demographic inference \citep{Ragsdale2019-nt}.  We
    infer detailed demographic models for populations across Africa, including
    representatives from eastern and western groups and newly sequenced whole
    genomes from 44 Nama (Khoe-San) individuals. We infer a reticulated African
    population history in which contemporary population structure dates back to
    Marine Isotope Stage (MIS) 5. The earliest population divergence among
    contemporary populations occurs 120--135ka and was preceded by structure
    between two or more weakly differentiated ancestral \emph{Homo} populations
    connected by gene flow over hundreds of thousands of years. Such \emph{weakly
    structured stem} models better explain patterns of
    polymorphism that were previously attributed to contributions from archaic
    hominins in Africa \citep{Plagnol2006-lt,Hammer2011-bx,Hey2018-pw,Ragsdale2019-nt,
Lorente-Galdos2019-vz,Durvasula2020-td}. In contrast to models with archaic introgression, we
    predict fossil remains from coexisting ancestral populations should be
    genetically and morphologically similar, and that an inferred 1--4\% of genetic
    differentiation among contemporary human populations can be attributed to
    genetic drift between stem populations. We show that model misspecification
    explains variation in previous divergence time estimates and argue that
    studying a suite of models is key to robust inferences about deep history.

\end{abstract}

\section*{Introduction}

Decades of study of human genome variation have suggested a predominantly
tree-like model of recent population divergence from a single ancestral
population in Africa. It has been difficult to reconcile this finding with the
fossil and archaeological records of human occupation across the vast African
continent. For example, fossils such as those from the sites of Jebel Irhoud,
Morocco \citep{Hublin2017-cq}, Herto, Ethiopia \citep{White2003-bk} and Klasies
River, South Africa \citep{Deacon1995-rx} demonstrate that derived \emph{Homo
sapiens} anatomical features were found across the continent
$300-100$ka. Archaeological sites from the Middle Stone Age, of which some have
been associated with \emph{Homo sapiens}, are also widely distributed across
Africa. It is unclear whether these fossils and archaeological sites represent
populations which contributed to contemporary \emph{Homo sapiens} as population
precedents or were local ``dead-ends''. Recently, attempts to reconcile genetic
and paleoanthropological data include proposals for a Pan-African origin of
\emph{Homo sapiens} by which populations in many regions of the continent
contributed to the formation of \emph{Homo sapiens} beginning at least $300$ka
\citep{Stringer2016-mj,Scerri2018-nl,Scerri2019-xg}.

Genetic models have been hampered in their contribution to this discussion
because they primarily assume (or, at least, have been tested under) a
tree-like model of isolation-with-migration. Alternative theoretical scenarios
have been proposed, such as stepping stone models \citep{Arredondo2021-qa} or
population coalescence and fragmentation \citep{Scerri2019-xg}, but these
approaches are more challenging to interpret and fit to data.
However, new population genetic tools now allow for inference involving tens to
hundreds of genomes from multiple populations and greater complexity
\citep{Kamm2020-vn,Ragsdale2019-nt,Speidel2019-nj}. Inspired by evidence for
Neanderthal admixture with humans in Eurasia, several recent studies
have shown that introducing an archaic hominin ``ghost'' population contributing to 
African populations in the period surrounding the Out-of-Africa migration event
substantially improves the description of genetic data relative to
single-origin models, mostly in western Africa
\citep{Plagnol2006-lt,Hammer2011-bx,Hey2018-pw,Ragsdale2019-nt,
Lorente-Galdos2019-vz,Durvasula2020-td},
%Plagnol: YRI 
%Here we present a new method for addressing whether archaic human groups contributed to the modern gene pool (called ancient admixture),
%we find strong evidence for ancient admixture in both a European and a West African population
%
%Hammer: Mandenka, Biaka, and San. Title:
%"Genetic evidence for archaic admixture in Africa"
%Abstract:"allow us to infer that contemporary African populations contain a small proportion of genetic material (≈2%) that introgressed ≈35 kya from an archaic population that split from the ancestors of anatomically modern humans ≈700 kya."
%
%Hey: Strongest signal in Yoruba, Baka
% Abstract: "consistent with human/archaic divergence."
%
%Ragsdale: YRI
%abstract: We estimate that an unidentified, deeply diverged population admixed with modern humans within Africa both before and after the split of African and Eurasian populations, contributing 4 − 8% genetic ancestry to individuals in world-wide populations.
%
%Lorente Galdos: Khoesan, Mbuti, W. AFR
%title:"reveals archaic gene flow"
%
%Durvasula: West African
%title: "Recovering signals of ghost archaic introgression"
%Abstract:"Our analyses of site frequency spectra indicate that these populations derive 2 to 19% of their genetic ancestry from an archaic population that diverged before the split of Neanderthals and modern humans." 
but also in southern \cite{Hammer2011-bx,Lorente-Galdos2019-vz} 
and central African \cite{Hsieh2016-gk, Hammer2011-bx,Lorente-Galdos2019-vz,Hey2018-pw}
populations.
This has driven speculation about the geographic range of this ghost
population, possible links to specific fossils, and the possibility of
finding ancient DNA evidence\citep{Hsieh2016-gk}. However, these prior
studies share two weaknesses. First, they only contrast a single-origin model
with an archaic hominin admixture model, leaving out other plausible models
\citep{Henn2018-rf} (Figure~\ref{fig:proposed-models}).
Second, they
focus on a small subset of African diversity, either because of small sample
sizes (2-5 genomes) or because they rely on 1000 Genomes data which only
recruited populations of recent West African or Bantu-speaking ancestry
(Figure~\ref{fig:diversity}C). 
While ancient DNA from Eurasia has helped to clarify early human history
outside of Africa, there is no comparably ancient DNA to elucidate early history
in Africa \citep{Lipson2022-xf}.

We therefore aim to discriminate among a broader set of demographic
models by studying the genomes of contemporary populations. We take as our
starting point four classes of models (single population expansion, single
population expansion with regional persistence, archaic hominin admixture, and
multi-regional evolution, Figure~\ref{fig:proposed-models}), using 290
genomes of individuals from southern, eastern, and western Africa as well as Eurasia.
By including geographically and genetically diverse populations across Africa, we
infer demographic models that explain more features of genetic diversity
in more populations than previously reported. These analyses confirm the inadequacy of
tree-like models and provide an opportunity to directly evaluate a wide range
of alternative models.

\section*{Results}

We inferred detailed demographic histories using 4x-8x whole-genome sequencing
data for four diverse African populations, comprising the Nama (Khoe-San from
South Africa, newly presented here, see Supplementary Information (SI) section
1.2 for ethical and practical aspects of participant recruitment), Mende (from Sierra Leone, MSL from the
Phase 3 1000 Genomes Project \citep{1000_Genomes_Project_Consortium2015-zq}),
Gumuz (recent descendants of a hunter-gatherer group from Ethiopia
\citep{Gurdasani2015-qy,Gopalan2022-pw}), and eastern African agriculturalists
(Amhara and Oromo from Ethiopia \citep{Gurdasani2015-qy}). The Amhara and Oromo
populations, despite speaking distinct Afro-Asiatic languages, are highly
genetically similar \citep{Pagani2015-pz,Gopalan2022-pw} and thus the two
groups were combined for a larger sample size (Figure~\ref{fig:diversity}). We also
included the British (GBR) from the 1000 Genomes Project in our demographic
models as a representative source of back-to-Africa gene flow and recent
colonial admixture in South Africa. Finally, we used a high-coverage ancient
Neanderthal genome from Vindija Cave, Croatia \citep{Prufer2017-kk} to account
for gene flow from Neanderthals into non-Africans and gauge the relative
time depth of divergence, assuming Neanderthals diverged $550$ka from a common
stem. We computed one- and two-locus statistics whose expectation within and across 
populations can be computed efficiently and that are well suited for both low- and
high-coverage genomes \citep{Ragsdale2019-nt,Ragsdale2020-nz}. Using a
maximum-likelihood inference framework, we then fit to these statistics a
family of parameterized demographic models that involve population splits, size
changes, continuous and variable migration rates, and punctuated admixture
events, to learn about the nature of population structure over the past million
years.

\subsection*{A Late Pleistocene common ancestry}

We began with a model of geographic expansion from a single ancestral,
unstructured source followed by migration between populations, without allowing
for contribution from an African archaic hominin lineage or population
structure prior to the expansion (Figure~\ref{fig:proposed-models}A or
Figure~\ref{fig:proposed-models}D). As expected \citep{Ragsdale2019-nt}, this
first model was a poor fit to the data qualitatively
(Figure~\ref{fig:supp-single-origin-fits}) and quantitatively (log-likelihood
($LL$) $\approx-$189,300, Table~\ref{tab:supp-single-origin}). We next
explored a suite of parameterized models in which population structure predates
the differentiation of contemporary groups (SI
Section~\ref{sec:modelspec}). Depending on the parameters, these encompassed
models allowing for ancestral reticulation, such as
fragmentation-and-coalescence or meta-population models
(Figure~\ref{fig:proposed-models}B), archaic hominin admixture
(Figure~\ref{fig:proposed-models}C), and African multi-regionalism
(Figure~\ref{fig:proposed-models}D). The recent expansion and the African
multi-regional models (Figure~\ref{fig:proposed-models}A and D) have the same
topology, so interpretation of the model depends on the specified or inferred
divergence times.

Regardless of the model choice for early epochs, maximum-likelihood inference
of human demographic history for the last $150$ka was remarkably robust.  In a
reticulated model, we use ``divergence'' between populations to mean the time
of their most recent shared ancestry. The earliest divergence among
contemporary human populations differentiates the southern African Nama from
other African groups between $110-135$ka, with low to moderate levels of
subsequent gene flow (Table \ref{tab:migration-rates}). In none of the
high-likelihood models which we explored did the divergence between Nama and
other populations exceed $\sim140$ka. We conclude that geographic patterns of
contemporary \emph{Homo sapiens} population structure likely arise during MIS
5. Although we find evidence for earlier population structure in Africa (next
section), contemporary populations cannot be easily mapped onto the more
ancient `stem' groups as only a small proportion of drift between contemporary
populations can be attributed to drift between stems (SI Section \ref{sec:f4},
Figures~\ref{fig:predictions} and
\ref{fig:supp-f4s-single-origin}--\ref{fig:supp-f4s-merger-with-stem-migration}).

Given this consistency in inferred recent history and the numerical challenge
of optimizing a large number of parameters, we fixed several parameters related
to recent population history so as to focus on more ancient events (SI Section
\ref{sec:models}). These parameters were ones supported by multiple genetic and
archaeological studies \citep{Bergstrom2021-iw}.
Fixed parameters included the time of divergence
between western and eastern African populations, set to $60$ka, just prior to
the split of Eurasians and East Africans set to $50$ka. We also fixed the
amount of admixture from Neanderthals to Europeans directly following the
out-of-Africa migration to $1.5\%$ at $45$ka.

We quantify migration rates among populations after their divergence
$\sim120$ka. Prior to agripastoralist expansion $5$ka, migration between the
ancestors of the Nama and other groups is an order of magnitude weaker than
that observed between western and eastern Africans (Table
\ref{tab:migration-rates}). All models infer relatively high gene flow between
eastern and western Africa ($m\approx2\times10^{-4}$, the constant proportion
of immigrant lineages per generation over $60$ka). We further find that
Back-to-Africa gene flow at the beginning of the Holocene primarily affected
the ancestors of the Ethiopian agricultural populations
\citep{Molinaro2019-rf}, comprising over half of their genetic ancestry,
estimated to be approximately $65\%$. We observe significant gene flow from the
Amhara and Oromo into the Nama, a signal which is likely a proxy for the
movement of eastern African caprid and cattle pastoralists
\citep{Henn2008-xo,Breton2014-xb}, here estimated to constitute a $25\%$ ancestry
contribution 2,000 years ago. While this gene flow is not apparent from the
\texttt{ADMIXTURE} plot (Figure~\ref{fig:diversity}), the ancestry is likely grouped
into the Khoe-San component which has drifted appreciably from its ancestral
eastern African source. Colonial period admixture from Europeans into the Nama
was estimated at $15\%$, similar to proportions suggested by \texttt{ADMIXTURE}
(Figure~\ref{fig:diversity}).

\begin{table}[t!]
    \centering
    \scriptsize
    \begin{tabular}{lllrrrr}
        \toprule
        \textbf{Likelihood} & \textbf{Label} & \textbf{Population Pair} &
            \textbf{Divergence} & \textbf{Migration rate} & \textbf{Migration} \\
        & & & 
            \textbf{Time (ka)} & \textbf{per generation} & \textbf{duration (kyr)} \\
        \midrule
        \textbf{Continuous Model} & & & & & \\
        $LL= -115,300$ & a & Stem 1, Stem 2 & 1,223 & 6.26e-5 & 1,089 \\
        (Table~\ref{tab:supp-continuous-migration}) & b & Stem 2, Nama & NA & 5.85e-5 & 129 \\
        & c, d & Stem 2, Other Africans* & NA & 3.10e-5, \textbf{1.62e-4} & 129, 55 \\
        & e, f & Nama, Other Africans* & 135 & 4.1e-5, 9.2e-6 & 134, 60 \\
        & g & Mende, East Africans & 60 & \textbf{2.13e-4} & 60 \\
        & h & East Africans, British & 50 & 4.16e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo & 12 &\textbf{3.37e-4} & 12 \\
        \textbf{Merger Model} & & & & & \\
        $LL= -101,600$ & a & Stem 1, Stem 2 & 1,692 & \textbf{1.26e-4} & 1,213 \\
        (Table~\ref{tab:supp-merger-with-stem-migration}) & -- & Stem 1S, Stem 1E & 479 & 0 (fixed) & -- \\
        & b & Stem 2 to Nama & 119 & \textbf{0.70} & pulse \\
        & c & Stem 2 to Stem 1E & 98 & \textbf{0.52} & pulse \\
        & d & Stem 2 to Mende & 25 & \textbf{0.19} & pulse \\
        & e, f & Nama, Other Africans* & 119 & 4.5e-5, 9.8e-6 & 120, 60 \\
        & g & Mende, East Africans* & 60 & \textbf{1.97e-4} & 60 \\
        & h & East Africans, British & 50 & 3.82e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo &12 & \textbf{3.59e-4} & 12 \\
        \bottomrule
    \end{tabular}
    \caption{
        \textbf{Migration and divergence parameters from best fit models.}
        Labeled migration rates correspond to symmetric continuous migration
        bands shown in Figure~\ref{fig:best-fit-models}. Both the continuous migration and
        merger models inferred a relatively deep split of human stem branches,
        though they were connected by ongoing migration that maintained their
        genetic similarity. Bold text indicates migration rates above $10^{-4}$.
        In both models, the branch ancestral to the Nama shares a common ancestral population
        with the other African groups $\sim120-135$ka. Following this divergence,
        the population ancestral to other African groups branches into West and East African
        groups $60$ka.
        *Migration rates and durations are shown between branches ancestral to 
        1) Nama and East Africans and their ancestors, and
        2) Nama and Mende, respectively.
        ``Divergence times'' correspond to the most recent common ancestral population
        and does not account for continuous migration or earlier reticulations.
    }
    \label{tab:migration-rates}
\end{table}

\subsection*{A weakly structured stem within Africa}

To account for population structure prior to $135$ka, three of our four models
allowed for two or more ``stem'' populations which could diverge either before
or after the Neanderthal split. We considered models both with and without
migration between these stem populations, and in both cases we tested two
different types of gene exchange during the expansion phase, as illustrated in
Figure \ref{fig:mergers}: 1) One of the stem population expands (splits into
contemporary populations), followed by continuous symmetric migration with
the other stem population(s); or 2) one or more of the stem
populations expands, with instantaneous pulse (or ``merger'') events from the
other stem population, so that recent populations are formed by mergers of
multiple ancestral populations. Depending on parameter values, this scenario
encompasses archaic hominin introgression and fragmentation-and-coalescence
models (such as Figure~\ref{fig:proposed-models}B and C). For many parameters,
confidence intervals based on bootstrapping are relatively narrow
(Tables~\ref{tab:supp-single-origin}--\ref{tab:supp-merger-with-stem-migration}),
reflecting an informative statistical approach. However, model assumptions have
a larger impact on parameter estimates (and thus real uncertainty).
To convey model uncertainty, we highlight features of the two inferred models
with high likelihoods. These are referred to as the ``multiple-merger'' and the
``continuous-migration'' models. Both allow for migration between stem
branches, but differ primarily in the timing of the early divergence of stem
populations and their relative $N_e$ (Figure~\ref{fig:best-fit-models}). The
two models also differ in the mode of divergence, with the multiple-merger
featuring a population reticulation during the Middle Pleistocene.

Allowing for continuous migration between the stem populations substantially
improved the fits relative to zero migration between stems ($LL \approx
-$101,600 vs.\ $-$107,700 in the merger model, Tables
\ref{tab:supp-merger-without-stem-migration} and
\ref{tab:supp-merger-with-stem-migration}, and $LL \approx -$115,300 vs.\
$-$126,500 in the continuous migration model, Tables
\ref{tab:supp-continuous-migration-no-mig} and
\ref{tab:supp-continuous-migration-no-mig}).  With continuous migration between
stems, population structure extends back to up to 1.5Ma
(Table~\ref{tab:migration-rates}).  Migration between the stems in these models
is moderate, with a fraction of migrant lineages each generation estimated as
$m=6.3\times10^{-5}-1.3\times10^{-4}$. For comparison, this is similar to
inferred migration rates between connected contemporary populations over the
past $50$ka (Table~\ref{tab:migration-rates}). This ongoing (or at least,
periodic) gene flow qualitatively distinguishes these models from previously
proposed archaic hominin admixture models (Figure~\ref{fig:proposed-models}C)
as the early branches remain closely related, and each branch contributes large
amounts to all contemporary populations (Figure~\ref{fig:predictions}).
Because of this relatedness, only $1\%$ to $4\%$ of genetic differentiation
among contemporary populations can be traced back to this early population
structure (SI Section~\ref{sec:f4}).

Under the continuous-migration model, one of the two stems diverges into
lineages leading to contemporary populations in western, southern and eastern
Africa, and the other (Stem 2) contributes variable ancestry to those
populations. This migration from Stem 2 is highest with the Mende
($m=1.6\times10^{-4}$) compared to the Nama and East African populations
($m=5.9\times10^{-5}$ and $3.1\times10^{-5}$, respectively), with migration
allowed to occur until $5$ka. A sampled lineage from the Nama, Mende, and Gumuz
have probabilities of being in Stem 2 at the time of Stem 1 expansion (135ka)
of approximately $0.145$, $0.2$, and $0.13$, respectively, though these
probabilities change over time, precluding the notion of a fixed admixture
proportion.

In contrast, under the multiple-merger model, stem populations merge with
varying proportions to form the different contemporary groups.  We observe a
sharp bottleneck in Stem 1 down to $N_e=100$ after the split of the Neanderthal
branch. This represents the lower bound allowed in our optimization (i.e., an
$N_e$ of 100), although the size of this bottleneck is poorly constrained (95\%
confidence interval 100--851). After a long period of exchange with Stem 2,
Stem 1 then fractures into ``Stem 1E'' and ``Stem 1S'' 478ka. The timing of
this divergence was also poorly constrained ($95\%$ CI $276-478$ka).  These
populations evolve independently until $119$ka ($101-125$ka) when Stem 1S and Stem
2 combine to form the ancestors of the Nama, with proportions 30\% and 70\%,
respectively. Similarly, Stem 1E and Stem 2 combine in equal proportions (50\%
each) to form the ancestors of the western Africans and eastern Africans (and
thus also all individuals who later disperse during the Out-of-Africa event).
Finally, the Mende receive a large additional pulse of gene flow from Stem 2,
replacing 19\% (18--21\%) of their population $25$ka ($22-26$ka). The later Stem 2
contribution to the western African Mende resulted in better model fits
($\Delta LL \approx 60,000$). Speculatively, this may indicate that an
ancestral Stem 2 population occupied western or central Africa, broadly
speaking. The differing proportions in the Nama and eastern Africans may also
indicate geographic separation of Stem 1S in southern Africa and Stem 1E in
eastern Africa.

To assess robustness of the inferred models to analysis and reference
population choices, SI Sections \ref{sec:validation_strategy}
and \ref{sec:validation_results} present re-analyses changing the European and
West African populations, as well as the recombination maps, filtering
strategies and parameter optimization strategies. While we find some
differences in the inferred parameters (see SI sections
\ref{sec:replacing} and \ref{sec:change-dates}), the best fit models across all
reanalyses are quantitatively consistent.

\subsection*{Reconciling lines of genetic evidence}

Previous studies have found support for archaic hominin admixture in Africa
using two-locus statistics \citep{Hsieh2016-gk,Ragsdale2019-nt}, the
conditional SFS (cSFS) \citep{Durvasula2020-td}, and reconstruction of gene
genealogies \citep{Speidel2019-nj}. However, none of these studies considered a
weakly structured stem. We validated our inferred models with additional
independent approaches. We find that the observed cSFS (conditioned on the
derived allele being carried in the Neanderthal sample) is very well-described
by the merger model (Figures~\ref{fig:validation}A-C and
\ref{fig:supp-csfs-single-origin}--\ref{fig:supp-csfs-merger-with-stem-migration}),
even though this statistic was not used in the fit. Our best-fit models
outperform archaic hominin admixture models fit directly to the cSFS (for
example, compare to Figure~1 in Durvasula and Sankararaman
(2020)\citep{Durvasula2020-td}). Specifically, it is the addition of migration
between stems that results in a qualitative improvement of the agreement
(compare Figures \ref{fig:supp-csfs-merger-without-stem-migration}  and
\ref{fig:supp-csfs-merger-with-stem-migration}).

We used \texttt{Relate} \citep{Speidel2019-nj} to infer the distribution of
coalescence rates over time in real data and data simulated from our inferred
models. Many previous studies have found a reduction of coalescence rates
between $1000$ka and $100$ka in humans, and thus inferred an increase in $N_e$
during the same period \citep{Li2011-le}.  This increase in inferred $N_e$
could be due to either an increase in population size or to ancestral
population structure during the Middle Pleistocene \citep{Mazet2016-wn}.  All
models, including the single-origin model, recapitulate an inferred ancestral
increase in $N_e$ between $100$ka--$1$Ma
(Figure~\ref{fig:supp-iicr-sim} and SI
Section~\ref{inferred_relate}). Whereas the single-origin model achieves this by an
increase in $N_e$ during that period, the best-fit models recapitulate this
pattern without corresponding population size changes.

Relative cross-coalescence rates (rCCR) have recently been used to estimate
divergence between pairs of populations, as measured by the rate of
coalescence between two groups divided by the mean within population
coalescence. Simulations of rCCR accuracy, however, focus on a ‘clean split’
between populations whereby groups diverge without subsequent gene flow.
Published estimates of the earliest human divergences with rCCR, which range
from $150$ka--$100$ka \citep{Bergstrom2021-iw}, may be significantly biased when
compared to more complex models with gene flow as inferred here. We find that
midpoint estimates of rCCR are poor estimates for population divergence, often
underestimating divergence time by 50\% or greater (e.g., Mende vs.\ Gumuz
$\sim15$ka compared to a true divergence of $60$ka), and recent migration can
lead to the misordering of divergence events (Figure~\ref{fig:validation}E). We
suggest that rCCR analyses which do not fit multiple parameters including gene
flow should be interpreted with caution.

Other studies have fit tree-like demographic models to African populations
using distributions of allele frequencies or related statistics, finding
inconsistent divergence times, some of which are older than those we find here
\cite{Henn2018-rf,Bergstrom2021-iw}. In the SI (Section~\ref{sec:IM-reinfer}),
we show that this discrepancy can be explained by model misspecification: if
divergence is estimated using an isolation with migration (IM) model with
constant population sizes, but the correct model has ancient population growth
or population structure, the divergence time in the inferred IM model is much
deeper than in the correct model. Intuitively, growth or structure in the
ancestral population will each increase coalescence times relative to a
randomly mating population of constant size, so a model that assumes constant
population sizes would require an older divergence time to fit the observed
distribution of coalescence times and related statistics
\citep{Momigliano2021-th,Shchur2022-wa}.

\section*{Discussion}

Any attempt at building detailed models of human history is subject to model
misspecification. This is true of earlier studies, which often assumed that
data inconsistent with a single-origin model must be explained by archaic
hominin admixture. This is also true of this study. While it remains
prohibitive to fully explore the space of plausible models of early human
population structure, we sought to capture model uncertainty by exploring
multiple parameterizations of early history. The best fit models presented
here include reticulation and migration between early human populations rather
than archaic hominin admixture from long-isolated branches
(Figure~\ref{fig:proposed-models}C). 
Elements of a recent single-origin or African multi-regionalism
(Figure~\ref{fig:proposed-models}A,D)
feature in our best fit models, as indicated in the recent time of
contemporary population divergence
or the gene flow between disparate stems, respectively.

We caution that we cannot rule out that more complex models involving
additional stems, more complex population structure, or
hybrid models including both weak structure and archaic hominin admixture may
better explain the data. Because parameters related to the split time,
migration rates, and relative sizes of the early stems were variable across
models, reflecting a degree of confounding among these parameters, we refrained
from introducing additional branches associated with more parameters during
that period. Rather than interpreting the two stems as representing
well-defined and stable populations over hundreds of thousands of years, we
interpret the weakly structured stem as consistent with a population
coalescence and fragmentation model \citep{Scerri2019-xg}.
Additional African populations such as those 
from Central Africa, other Khoe-San groups or pre-Holocene ancient DNA samples,
could further test our proposed models.

\subsection*{Formation of Population Structure in Africa}

Our inferred models paint a consistent picture of the Middle to Late
Pleistocene as a critical period of change, assuming that estimates from the
recombination clock accurately relate to geological chronologies (SI
Section~\ref{sec:supp-mutation}). During the late Middle Pleistocene, the
multiple merger model indicates three major stem lineages in Africa,
tentatively assigned to southern (Stem 1S), eastern (Stem 1E) and
western/central Africa (Stem 2). Geographic association was informed by the
present population location with the greatest ancestry contribution from each
stem. For example, Stem 1S contributes 70\% to the ancestral formation of the
Khoe-San. The extent of the isolation $\>400$ka between Stem 1S, Stem 1E and Stem
2 suggests that these stems were not proximate to each other. While the length
of isolation among the stems is variable across fits, models with a period of
divergence, isolation and then a merger event (i.e., a ``reticulation'')
out-performed models with bifurcating divergence and continuous gene flow.

A population reticulation involves multiple stems contributing genetically to
the formation of a group. One way in which this can happen is through the
geographic expansion of one or both stems. For example, if during MIS 5, either
Stem 1S (Figure~\ref{fig:best-fit-models}B) from southern Africa moved
northward thereby encountering Stem 2, or Stem 2 moved from central/western
Africa southward into Stem 1S, then we could observe disproportionate
ancestry contributions from different stems in contemporary groups. We observed
two merger events. The first, between Stem 1S and Stem 2, resulted in the
formation of an ancestral Khoe-San population $\sim120$ka. The second event
$\sim100$ka between Stem 1E and Stem 2, resulted in the formation of the
ancestors of eastern/western Africans including the ancestors of
``Out-of-Africans''. Reticulated models do not have a unique and 
well-defined ``basal'' human population divergence.
We suggest conceptualizing the events $120$ka as the time of most recent
shared ancestry among sampled populations.
However, interpreting population divergence times in population genetics is 
always challenging due
to the co-estimation of divergence time and subsequent migration;
methods assuming clean vs reticulated splits can infer different split dates
(Figure~\ref{fig:supp-rccr-data} and Figure~\ref{fig:supp-IM-misspecification}).
Therefore, wide variation exists in estimates of divergence time in the
literature\citep{Bergstrom2021-iw,Henn2018-rf}.

Shifts in wet/dry conditions across the African continent $140$ka to $100$ka may have
promoted these merger events between divergent stems. 
Precipitation does not neatly track interglacial cycles in Africa,
and heterogeneity across regions may mean that the beginning of an arid period
in eastern Africa is conversely the start of a wet period in
southern Africa\citep{Blome2012-lw}.
The rapid rise in sea levels during the MIS 5e interglacial might have
triggered migration inland away from the
coasts, as has been suggested, e.g., for the Paleo-Agulhas plain\citep{Marean2014-pg}.
Following these merger events, the stems subsequently fractured into
subpopulations which then persist over the past $\sim120$ka. These
subpopulations can be linked to contemporary groups despite subsequent gene
flow across the continent; for example, a genetic lineage sampled in the Gumuz
has a $0.70$ probability of being inherited from the ancestral ``eastern''
subpopulation $55$ka versus $0.06$ probability of being inherited
from the ``southern'' subpopulation (see Table~\ref{tab:supp-lineage-probabilitys} for
additional comparisons).

We also find that Stem 2 continued to contribute to western Africans during the
Last Glacial Period, indicative that this gene flow likely occurred in
western/central Africa (Table~\ref{tab:migration-rates}). Such an
interpretation is reinforced by the differential migration rates between
regions, i.e., the gene flow from Stem 2 to western Africans is estimated to be
five times that of the rate to eastern Africans during this period. We
performed a variety of validation tests to explore sensitivity of our
assumptions, including relaxing fixed parameters (SI
Section~\ref{sec:validation_strategy}). Most validation tests resulted in
parameters similar to the models discussed above. However, one exception was
the inferred Out-of-Africa and eastern/western African divergences which were
10--15ka younger than our fixed parameters. These younger dates are clearly at 
odds with the accepted timing of Out-of-Africa expansion at approximately $\sim>50$ka, 
as based on archaeological, climatic and paleo-morphological information
\citep{Fu2014-am,Groucutt2015-do,Prufer2021-ao,Beyer2021-ym}. 
We interpret the free estimate for eastern African vs. European divergence as reflecting 
our inclusion of only a single OOA population in the model, the lack of a nearby 
source for back-to-Africa gene flow, and other regionally complex parameters. 
However, older pan-African features of our inferred models are weakly affected by 
this analysis choice (SI Section~\ref{sec:models}).

\subsection*{Contrasting ancestral structure models}

Evidence for archaic hominin admixture in Eurasia has bolstered the
plausibility of archaic hominin admixture having also occurred in Africa.
Previous work that sought to explain patterns of polymorphism inconsistent with
a single-origin model thus focused on archaic hominin admixture as an
alternative model. The additional (or ``ghost'') branches required to fit the
data are systematically referred to as ``archaic''
\citep{Plagnol2006-lt,Hammer2011-bx,Hey2018-pw,Ragsdale2019-nt,
Lorente-Galdos2019-vz,Durvasula2020-td}, and assumed (or inferred) to be deeply
diverged. We believe that this has strongly oriented the interpretation of
these results (e.g., regarding selection \cite{Wall2019-ao} or the fossil
record \cite{Grun2020-su}). Here, we have shown that a model of a weakly
structured stem better captures the apparently inconsistent patterns of
polymorphisms.

Preferring weakly-structured-stem models over archaic-admixture models
potentially has a range of implications. First, with a weakly structured stem,
there is no need to posit that an archaic hominin population in Africa stayed
reproductively isolated from the ancestral human lineage for hundreds of thousands of
years before the initiation of gene flow. Instead, there would simply have
been continuous or recurrent contact between two or more groups present in
Africa.

Second, there is evidence for both deleterious and adaptive
archaic-hominin-derived alleles in contemporary genomes in the form of a
depletion of Neanderthal ancestry in regulatory regions \citep{Petr2019-xo} or
an increased frequency of archaic-hominin-related haplotypes such as at
\emph{EPAS1} among Tibetans \citep{Zhang2021-xx}. Under previous
African-archaic-hominin admixture models, the estimated 8--10\% introgression rate is
much higher than Neanderthal gene flow, and would have plausibly been fertile
ground for dramatic selection for or against archaic-hominin-derived
haplotypes\citep{Wall2019-ao}. By contrast, adaptation under a weakly
structured stem would have occurred continuously over much longer periods.
Patterns of polymorphism that are inconsistent with the single-stem model
predictions have been used to infer putative archaic admixed segments
\citep{Plagnol2006-lt,Hsieh2016-gk,Wall2019-ao,Durvasula2020-td}, negative
selection against such segments \citep{Wall2019-ao}, and pervasive positive
selection \citep{Schrider2017-kl}. However, such approaches are subject to high
false positives in the presence of population structure with migration
\citep{Petr2019-xo}, and their interpretation should be re-examined in light of
a weakly-structured-stem model within Africa.

Third, multiple studies have shown a correspondence between phenotypic
differentiation, usually assessed with measurements of the cranium, and genetic
differentiation among human populations and between humans and Neanderthals
\citep{Relethford1994-mh,Weaver2008-ho,Von_Cramon-Taubadel2009-zb} (see also SI
Section~\ref{sec:morpho_diff}). This correspondence potentially allows
predictions of our model to be related to the fossil record. The fossil record
of Africa is sparse during the time period of the stems, but of the fossils
that date to this time period, some are fairly similar overall in morphology to
contemporary humans (e.g., Omo 1 from Omo Kibish, Ethiopia
\citep{Day1969-rh,Vidal2022-qe}), others are similar in some morphological
features to contemporary humans (e.g., Irhoud 1 from Jebel Irhoud, Morocco
\citep{Hublin2017-cq,Richter2017-zu}), and others are different enough in
morphology to have been assigned to species other than \emph{Homo sapiens}
(e.g., DH1 from Dinaledi, South Africa \citep{Berger2015-bq,Dirks2017-uk}). If,
as our model predicts, the genetic differences between the stems were
comparable to those among contemporary human populations, the most
morphologically divergent fossils are unlikely to represent branches that
contributed appreciably to contemporary human ancestries.

\bibliographystyle{naturemag}
\bibliography{paper}

\section*{Supplementary information}
Available in the online version of the paper.

\section*{Acknowledgements}

We are grateful for the DNA contribution from each participant that enabled
this study; in particular we wish to highlight the generous participation of
the Richtersveld Nama community in South Africa and help from local research
assistants Willem DeKlerk and Hendrik Kaimann. Additional assistance and
community engagement was conducted by Justin Myrick, Chris Gignoux, Caitlen
Uren and Cedric Werely. We thank the African Genome Diversity Project for data
generation, including Tommy Carensten, Deepti Gurdasani, and Manj Sandhu. We
thank Luke Anderson-Trocm\'e and Georgette Femerling for assistance in creating
the map in Figure~\ref{fig:diversity} and Figure \ref{fig:toymodels},
respectively. We thank Nuria Melisa Morales-Garcia for extensive data
visualization discussion and designing Figures \ref{fig:proposed-models} and
\ref{fig:best-fit-models}. This research was supported by CIHR project grant 437576,
Natural Sciences and Engineering Research Council of Canada (NSERC) grant RGPIN-2017-04816, 
and the Canada Research Chair program. 
This research was supported by an NIH grant
R35GM133531 (to BMH). The content is solely the responsibility of the authors
and does not necessarily represent the official views of the National
Institutes of Health.

\section*{Competing interests}
The authors declare no competing financial interests.

\section*{Author contributions}

\sgcomment{Rough first pass for contributions:}
APR, BMH, and SG designed the study. BMH, EGA, EH, and MM designed and conducted 
participant recruitment.   APR, BMH, and SG designed statistical analyses.
APR conducted statistical analyses. 
APR, TDW, BMH, and SG interpreted the results and wrote the first draft of the article. 
All authors read and edited the article.

\section*{Methods}

\subsection*{Data and sequencing}

We used a subset of populations from the merged sequencing data from the
African Diversity Reference Panel (ADRP) \citep{Gurdasani2015-qy,Pagani2015-pz}
and the Thousand Genomes Project (1KGP)
\citep{1000_Genomes_Project_Consortium2015-zq}. These included the Amhara,
Bakiga, Gumuz, Nama, Oromo, Somali, and Zulu from ADRP, and ESN, GWD, LWK, MSL,
YRI, CEU, GBR, CHB, and PJL from 1KGP. After filtering for relatedness and
retaining Nama individuals with $>70\%$ estimated Khoe-San ancestry, we used
data from 289 individuals, including 44 Nama individuals. These were merged
with the high-coverage Neanderthal genome from Vindija Cave
\citep{Prufer2017-kk}. We kept variants from regions that fell within the 1KGP
srict callability mask, overlapped with at least 100 continuously called base
pairs in the Neanderthal genome, and were annotated as intergenic.
\texttt{ADMIXTURE} and PCA analyses were conducted on a subset of variants
filtered to remove variants in high LD ($r^2$ threshold of $0.1$). See SI
section \ref{sec:data-sequencing} for additional details on data and
sequencing.

\subsection*{LD and diversity statistics}

We used multi-population LD and pairwise diversity statistics to fit
parameterized demographic models to data \citep{Ragsdale2019-nt}. Unbiased LD
statistics were computed from all variants in retained intergenic regions
\citep{Ragsdale2020-nz}, for pairs of variants separated by recombination
distances $r = 5\times10^{-6}$ to $5\times10^{-3}$. These were assigned to 16
recombination distance bins, and average statistics were computed within each
bin (SI section \ref{sec:computing-stats}). Expected statistics were computed
in \texttt{moments}, which also performed likelihood-based parameter
optimization. Conditional site frequency spectra (cSFS) were computed
conditioned on the Vindija Neanderthal carrying the derived allele relative to
the ancestral allele determined by a six-primate alignment
\citep{1000_Genomes_Project_Consortium2015-zq}.

\subsection*{Model specification and fitting}

Model parameters include population sizes and size changes, split times,
continuous migration rates, and admixture times and proportions. The simplest
model we tested was a bifurcating tree-like structure, allowing for subsequent
migrations and recent known admixture events. To include ancestral population
structure, we tested models that included multiple stem groups, which each were
allowed their own population size and could be connected by continuous
migration. We tested multiple scenarios of early population structure,
including long-lasting continuous migration between stem populations and
scenarios of periods of isolation with subsequent merger events (SI
section~\ref{sec:modelspec}). To avoid overfitting, we to incrementally added
copmlexity to our model optimization, and we fixed a number of parameters that
are constrained by historical records or are consistently estimated across
multiple models and previous studies (SI section~\ref{sec:models}). Likelihoods
were computed using a composite multivariate Gaussian likelihood approach, and
confidence intervals were estimated by refitting each model to 200
block-bootstrap replicate datasets (SI
section~\ref{sec:likelihoods-uncertainty}). We iteratively used gradient
descent and L-BFGS-B optimization routines to fit each parameterized model (SI
section~\ref{sec:optimization}).

\subsection*{Gene genalogy reconstruction}

\subsection*{Conditional site frequency spectrum}

\section*{Data availability}

Nama sequencing data are available from the European Genome-Phenome Archive (EGA),
accession number: EGAD00001006198. The African Diversity Reference Panel can be
found at accession: EGAS00001000960.

\section*{Code availability}

Code for the software used in this paper is found at the following locations:\\
moments-LD (\url{https://bitbucket.org/simongravel/moments}),\\
Demes (\url{https://github.com/popsim-consortium/demes-python}),\\
Relate (\url{https://myersgroup.github.io/relate/}),\\
msprime (\url{https://github.com/tskit-dev/msprime}),\\
tskit (\url{https://github.com/tskit-dev/tskit}),\\
and scripts implementing analyses using each of these software are available at\\
\url{https://github.com/apragsdale/african-structure-paper}.

\break

\section*{Figure captions}

\begin{figure}[ht]
    \centering
    %\includegraphics[width=0.5\textwidth]{figures/proposed-models.png}
    \caption{
        \textbf{Proposed conceptual models of early human history in Africa.}
        These models have been designed to translate models from the 
        paleoanthropological literature into genetically testable demographic
        models \citep{Henn2018-rf}.
        We used these conceptual models as starting points to build detailed
        parameterized demographic models (SI Section~\ref{sec:modelspec})
        that were then fit to genetic data. 
    }
    \label{fig:proposed-models}
\end{figure}

\begin{figure}[ht!]
    \centering
    %\includegraphics{figures/samples-diversity.pdf}
    \caption[width=\textwidth]{
        \textbf{Genetic diversity across Africa.}
        (A) Select populations from the 1000 Genomes and African Diversity Reference Panels
        illustrate diversity from western, eastern and southern Africa.
        We chose representatives from each region (bold labels)
        to build parameterized models,
        including the newly-sequenced Nama from South Africa, Mende
        from Sierre Leone, Gumuz, Oromo and Amhara from Ethiopia, and the
        British and Vindija Neanderthal individual.
        (B, C) PCA highlights the range of genetic divergence anchored 
        by western Africans, Nama, Gumuz and the British.
        Percentages show variance explained by each principal component.
        (D) \texttt{ADMIXTURE} with $K=4$ illustrates signatures of recent gene flow in Africa
        which reflect colonial-period migration into the Nama,
        Back-to-Africa gene flow among some Ethiopians, and Khoe-San admixture in the Zulu.
    }
    \label{fig:diversity}
\end{figure}

\begin{figure}[t!]
    \centering
    %\includegraphics[width=0.65\textwidth]{figures/best-fit-models}
    \caption{
        \textbf{A weakly structured stem best describes two-locus statistics.}
        In the two best-fitting parameterizations of early population structure,
        continuous migration (A) and multiple mergers (B), models
        that include ongoing migration between stem populations outperform
        those in which stem populations are isolated. Most recent populations are
        connected by continuous, reciprocal migration that are indicated by 
        double-headed arrows (labels matched to migration rates and divergence
        times in Table~\ref{tab:migration-rates}). These migrations last for the
        duration of co-existence of contemporaneous populations with constant
        migration rates over those intervals. The
        merger-with-stem-migration model (B, with  $LL=-101,600$) outperformed the
        continuous-migration model (A, with $LL=-115,300$).
        Colors are used to distinguish overlapping branches and link to
        Figure~\ref{fig:diversity}.}
    \label{fig:best-fit-models}
\end{figure}

\begin{figure}[t!]
    \centering
    %\includegraphics{figures/predictions.pdf}
    \caption{
        \textbf{Structure among stems is weak and present-day structure is mostly recent.}
        From the best fit models of our two parameterizations (A and B:
        continuous migration, C and D: merger with stem migration),
        we predicted pairwise differences $H_{i,j}$ between individuals
        sampled from populations $i$ and $j$ existing at time $t$ (A and C).
        (B and D) To understand how drift between stems explains contemporary structure, 
        we also computed the proportion $\alpha^2$ of drift between pairs of
        sampled contemporary
        populations that aligns with drift between past populations
        (here Nama and Mende, see SI Section~S\ref{sec:f4}
        for details and additional comparisons in
        Figures~\ref{fig:supp-f4s-single-origin}--\ref{fig:supp-f4s-merger-with-stem-migration}).
        Both models infer deep population structure with modest contributions to
        contemporary genetic differentiation.
        Most present differentiation dates back to the last $100$ka.
    }
    \label{fig:predictions}
\end{figure}

\begin{figure}[t!]
    \centering
    %\includegraphics{figures/validation}
    \caption{
        \textbf{Model validation using independent statistics.} (A--C) Using
        our best fit models to linkage disequilibrium and pairwise diversity statistics,
        we simulated expected conditional site-frequency-spectra (cSFS) and
        compared to the observed cSFS from the data. Our inferred models
        provide a good fit to the data, even though these were not used in our
        inference. Across the three populations, ancestral state
        misidentification was consistently inferred to be $1.5-1.7\%$ for
        intergenic loci (SI \ref{sec:misspecification}). (D, E) We used Relate
        \citep{Speidel2019-nj} to reconstruct genome-wide gene genealogies,
        which we used to estimate coalescence rate trajectories
        and cross-coalescence rates between pairs of populations. While
        coalescence rate distributions are informative statistics about past
        evolutionary processes, interpretation can be hindered by migration and
        population structure, and translating relative cross-coalescence rate
        curves (RCCR) into population divergence times is especially prone
        to misinterpretation. For example, the Mende-Gumuz comparison shows
        a more recent increased RCCR than either population with the British,
        a pattern that is recapitulated under our best-fit model, even though
        the Mende-Gumuz split occurs prior to the Gumuz-British split.
    }
    \label{fig:validation}
\end{figure}

\end{document}